---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# d2r

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

The goal of d2r is to support the creation, reading, writing, and rendering of D2 diagrams using R.

What is D2? From the [D2 Introduction](https://d2lang.com/tour/intro):

> D2 is a diagram scripting language that turns text to diagrams. It stands for Declarative Diagramming. Declarative, as in, you describe what you want diagrammed, it generates the image.

I am a contributor to the [D2 Extension for Quarto](https://github.com/data-intuitive/quarto-d2) and created this package to experiment with the possibilities of native R diagram creation and rendering.

Note that this package is experimental and the process for building and editing D2 diagrams is likely to change. I also hope to add a knitr engine to support d2 code blocks within a RMarkdown or Quarto document.

## Installation

You can install the development version of d2r like so:

``` r
# pak::pkg_install("elipousson/d2r")
```

## Example

```{r example}
library(d2r)
## basic example code
```

If d2 is installed and available on your PATH, you can check the version:

```{r}
d2_version()
```

If you have D2 installed with Homebrew, you may have trouble rendering a diagram within a RMarkdown. If so, you may need to set an environmental variable that points to the D2 PATH.

Use `d2_which()` to locate the path for the installed version of D2:

```{r}
d2_which()
```

This [2018 blog post by Tony Tsai](https://blog.tonytsai.name/blog/2018-05-07-setting-path-variable-for-gs-command-in-rstudio/) on trouble-shooting `PATH` variable issues for `brew`-installed commands.

Here is an example of basic diagram creation:

```{r shapes}
d2_diagram(c("a" = "b"))
```

Example of diagram with specified connectors:

```{r connectors}
d2_diagram(
  c("x" = "y", "y" = "z"),
  connector = c("->", "<-")
)
```

Example of rendered diagram:

```{r}
diagram <- d2_diagram(c("a" = "b", "x" = "y"))  

file <- tempfile(fileext = ".d2")

output <- paste0(knitr::opts_chunk$get("fig.path"), "example.png")

d2_write(diagram, file)

d2_render(file, output, theme = "everglade green")

knitr::include_graphics(output)
```

